---
title: "Assignment1"
author: "Junhao Ou & 21200066"
date: "2022-09-26"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

# Task 1: Manipulation

## 1. Load the dataset EurostatCrime2019.csv

In this section, we will load the dataset

```{r}
df = read.csv('EurostatCrime2019.csv',row.names = 1)
```

## 2. Structure Check

In this section, we will check the size (number of rows and columns) and the structure of this data.

```{r}
str(df)
```

As we can see, the size of the table is 41 rows by 13 columns.

## 3. Produce appropriate commands to do the following actions.

### (i) Remove the columns of Rape and Sexual.assault.

```{r}
df = df[,-which(names(df) %in% c('Rape','Sexual.assault'))]
str(df)
```

### (ii) Remove the columns involving theft and burglary
```{r}
library(dplyr)
df = select(df, -contains("Theft"))
df = select(df, -contains("Burglary"))
str(df)
```

### (iii) Add a column containing the overall record of offences for each country (per hundred thousand inhabitants)
```{r}
df['Overall Record'] = rowSums(df[,c(1:7)])
```

## 4. List the countries that contain any missing data.
```{r}
row.names(df[rowSums(is.na(df)) > 0,])
```

## 5. Remove the countries with missing data from the dataframe.
```{r}
df = subset(df,rowSums(is.na(df)) <= 0)
```

## 6. How many observations and variables are in this new dataframe?
```{r}
str(df)
```

# Task 2: Analysis

## 1. According to these data what were the 3 most common crimes in Ireland in 2019?

Then, we will implement a function that can return the 3 most common crimes with input country.

```{r}
maxthree = function(data, row) {
    x <- data[row,1:7] 
    return(x[order(x, decreasing = T)][1:3])
}
```

Check the 3 most common crimes in Ireland in 2019

```{r}
maxthree(df, "Ireland")
```

## 2. What proportion of the overall crimes was due to Assault in Ireland in 2019?

We will implement a function that can return the proportion of any type of crime of any country.
```{r}
propCal = function(data,country,crime) {
  prop = data[country,crime]/data[country,'Overall Record']
  return(prop)
}
```

Check the proportion of the overall crimes was due to Assault in Ireland.
```{r}
propCal(df, "Ireland",'Assault')
```

## 3. Which country had the highest record of kidnapping in 2019.
```{r}
rownames(df)[order(df[,'Kidnapping'], decreasing = T)][1]
```

## 4. Which country had the lowest overall record of offences in 2019 (per hundred thousand inhabitants)?
```{r}
rownames(df)[order(df[,'Overall Record'], decreasing = F)][1]
```

## 5. Create a plot displaying the relationship between robbery and unlawful acts involving controlled drugs or precursors. Make the plot look “nice” i.e. change axis labels etc.

```{r}
plot(df$Robbery,df$Unlawful.acts.involving.controlled.drugs.or.precursors,xlab='Robbery',ylab='Unlawful acts',main='The relationship between robbery and unlawful acts')
```

# Task 3: Creativity
1. We will plot the histogram of Overall Record for each country
```{r}
library(reshape2)
library(ggplot2)
df1 <- melt(df$'Overall Record',id.vars="Row.names")
ggplot(df1, aes(x = rownames(df), y=value))+geom_bar(stat="identity", position ="dodge")+theme_bw()+theme(axis.text.x = element_text(angle=-40, hjust=.1))+labs(x = "Country",y='Number of overall records')
```


From the plot above, we can conclude that the top 5 countries with the greatest number of crime records are Luxembourg, Swizerland, Finland, Germany and Ireland.